Budgeted Reinforcement Learning in Continuous State Space

### 2019-10-14 paper/supplementary.tex

This paper gathers all the supplementary material and goes as follows: \Cref{sec:proofs} details all the proofs of the main results. \Cref{sec:risk-sensitive-supp} and \Cref{sec:bftq-full} recall respectively the scalable \BFTQ algorithm and the risk-sensitive exploration procedure. \Cref{sec:lagragian} describes a naive alternative to \BFTQ based on Lagrangian Relaxation. The \Cref{sec:exp-supp} assembles all the assets for visualising and reproducing the experiments, including visualisations of policy executions, algorithms and environment parameters, and instructions for executing the attached source code. Finally we fill the Machine Learning Reproducibility Checklist and we justify each statement in \Cref{sec:ml-checklist}.

@@

\begin{proof}
This proof is the same as that in classical multi-objective MDPs.

@@

\begin{align}
    Q_r^*(\os, \oa) &=  \max_{\rho, \pi'} Q_r^{\rho, \pi'}(\os', \oa') \label{eq:pthm_def}\\
    &= \max_{\rho, \pi'} R_r(s, a) + \gamma \sum_{\os'\in\cS} P(\os' | \os, \oa) V_r^{\rho, \pi'}(\os') \label{eq:pthm_exp}\\
    &= R_r(s, a) + \gamma \sum_{\os'\in\cS}  P(\os' | \os, \oa) \max_{\rho, \pi'} \sum_{\oa'\in\ocA} \rho(\oa' | \os')Q_r^{\pi'}(\os', \oa') \label{eq:pthm_marg}\\
    &= R_r(s, a) + \gamma \sum_{\os'\in\cS}  P(\os' | \os, \oa) \max_\rho\sum_{\oa'\in\ocA}\rho(\oa' | \os')\max_{\pi'\in\Pi_a(\os')}Q_r^{\pi'}(\os', \oa') \label{eq:pthm_max}\\
    &= R_r(s, a) + \gamma \sum_{\os'\in\cS}  P(\os' | \os, \oa) \max_\rho\expectedvalueover{\oa'\sim\rho}Q_r^*(\os', \oa') \label{eq:pthm_marg_def2}
\end{align}
where $\pi = (\rho, \pi')\in\Pi_a(\os)$ and $\pi'\in\Pi_a(\os')$.

This follows from:
\begin{enumerate}
\item[\eqref{eq:pthm_def}.] Definition of $Q^*$. 
\item[\eqref{eq:pthm_exp}.] Bellman Expectation expansion from \Cref{prop:bellman-expectation}.
\item[\eqref{eq:pthm_marg}.] Marginalisation on $\oa'$.
\item[\eqref{eq:pthm_max}.] \begin{itemize}
    \item Trivially $\max_{\pi'\in\Pi_a(\os')} \sum_{\oa'\in\cA} \cdot \leq \sum_{\oa'\in\cA} \max_{ \pi'\in\Pi_a(\os)} \cdot$
    \item Let $\ov{\pi}\in\argmax_{\pi'\in\Pi_a(\os')} Q_r^{\pi'}(\os', \oa')$, then:
    \begin{align*}
        \sum_{\oa'\in\ov{A}}\rho(\oa'|\os')\max_{\pi'\in\Pi_a(\os')}Q_r^{\pi'}(\os', \oa') &= \sum_{\oa'\in\ov{A}}\rho(\oa'|\os')Q_r^{\ov{\pi}}(\os', \oa') \\
        &\leq  \max_{\pi'\in\Pi_a(\os')} \sum_{\oa'\in\ov{A}}\rho(\oa'|\os')Q_r^{\pi'}(\os', \oa')
    \end{align*}
\end{itemize}
\item[\eqref{eq:pthm_marg_def2}.] Definition of $Q^*$.
\end{enumerate}

@@

\subsection{\Cref{rmk:contractivity-smooth}}
\label{proof:contraction-with-smooth}
\begin{proof}
We now study the contractivity of $\cT$ when restricted to the functions of $\cL_\gamma$ defined as follows:
\begin{equation}
    \cL_\gamma = \left\{\begin{array}{cc}
         Q\in(\Real^2)^{\ocS\ocA}\text{ s.t. }\exists L<\frac{1}{\gamma}-1: \forall \os\in\ocS,\oa_1,\oa_2\in\ocA,   \\
         |Q_r(\os,\oa_1) - Q_r(\os,\oa_2)| \leq L|Q_c(\os,\oa_1) - Q_c(\os,\oa_2)|
    \end{array}\right\}
\end{equation}
That is, for all state $\os$, the set $Q(\os, \ocA)$ plot in the $(Q_c,Q_r)$ plane must be the \emph{graph} of a $L$-Lipschitz function, with $L<1/\gamma-1$.

@@

This sketch of proof makes use of insights detailed in the proof of \Cref{prop:bftq_pi_hull}, which we recommend the reader to consult first.

@@

\section{Risk-Sensitive Exploration}
\label{sec:risk-sensitive-supp}
We recall the Risk-Sensitive Exploration in \Cref{algo:risk-sensitive-exploration}:

\input{source/risk-sensitive-explo-pseudo-code.tex}

@@

\subsection{Examples of different exploration strategies}
\label{subsec:exploration-examples}
We compare two approaches for constructing a batch of samples. The animations from the html page \url{exploration.html} display the trajectories collected in each intermediate sub-batch. The first row corresponds to a classical  risk neutral epsilon-greedy exploration policy while the second row showcases a risk-sensitive exploration strategy introduced in the paper. Each animation corresponds to a different seed.

@@

On the \texttt{highway-env} , the budgeted agents display a wide variety of behaviours. Animations are displayed on the html page \url{highway-env.html}. When $\beta = 1$, the ego-vehicle drives in a very aggressive style: it immediately switches to the opposite lane and drives as fast as possible to pass slower vehicles, swiftly changing lanes to avoid incoming traffic. On the contrary when $\beta = 0$, the ego-vehicle is conservative: it stays on its lane and drives at a low velocity. With intermediate budgets such as $\beta = 0.2$, the agent sometimes decides to overtake its front vehicle but promptly steers back to its original lane afterwards.

@@

Animations are displayed on the html page \url{corridors.html} for the \texttt{corridors} environment. When the budget is low, the agent takes the safest path on the left. When the budget increases, it gradually switches to the other lane, earning higher rewards but also costs. This gradual process could not be achieved with a deterministic policy as it would chose either one path or the other. Each animation corresponds to a different seed.

@@

The following section displays environments and algorithms parameters and instructions to reproduce the exact same results displayed in \Cref{sec:experiements}.

@@

\subsubsection{Instructions for reproducibility}
\label{subsubsec:instruction-reproducibility}
To reproduce the result displayed in \Cref{sec:experiements}, first install the following conventional libraries for python3: pycairo, numpy, scipy and pytorch. Then, execute the commands in \Cref{fig:instructions} on a Linux Operating System. The Graphic Processing Unit used for experiments is an NVIDIA \texttt{GeForce GTX 1080 Ti} and the Computational Processing Unit is an Intel \texttt{Xeon E7}.

@@

%\begin{minted}{sh}
\begin{lstlisting}[language=bash,caption={bash version}]
# Install highway-env
pip3 install --user git+https://github.com/eleurent/rl-agents
# Change python path to the path of this repository
export PYTHONPATH="code/scaling-up-brl"
# Navigate to budgeted-rl folder
cd code/scaling-up-brl/budgeted-rl/
# Run main script using any config file
# Choose the range of seeds you want to test on
python3 main/egreedy/main-egreedy.py config/slot-filling.json 0 6
python3 main/egreedy/main-egreedy.py config/corridors.json 0 4
python3 main/egreedy/main-egreedy.py config/highway-easy.json 0 10
\end{lstlisting}
%\end{minted}

@@

    \begin{itemize}\item \textbf{yes}, see \Cref{sec:intro}, \Cref{sec:bdp}, \Cref{sec:brl},
    \Cref{sec:risk-sensitive-supp} and \Cref{sec:bftq-full}.
    \end{itemize}
    \item An analysis of the complexity (time, space, sample size) of any algorithm:
    \begin{itemize}\item \textbf{yes}, see \Cref{subsec:parallel-computing}.
    \end{itemize}
    \item A link to a downloadable source code, with specification of all dependencies, including external libraries:
    \begin{itemize}\item \textbf{yes}, see \Cref{subsubsec:instruction-reproducibility} and the folder \texttt{code} in the supplementary material zip file.
    \end{itemize}
\end{itemize}

@@

\begin{itemize}
    \item  A statement of the result:
     \begin{itemize}
        \item  \textbf{yes}, see \Cref{sec:bdp} and \Cref{sec:brl}.\end{itemize}
    \item A clear explanation of any assumptions:
    \begin{itemize}
        \item  we make one assumption in \Cref{sec:bdp}. We assume the program is feasible for any state. If not, no algorithm would be able to solve it anyway.\end{itemize}
    \item A complete proof of the claim:
    \begin{itemize}
        \item  \textbf{yes}, see \Cref{sec:proofs}. We formulate a conjecture in \Cref{rmk:contractivity-smooth} but we provide a sketch of the proof in \Cref{proof:contraction-with-smooth}.\end{itemize}
\end{itemize}

@@

        \begin{itemize}
            \item\textbf{yes}, see \Cref{sec:experiements} and \Cref{sec:algorithms-parameters}.
        \end{itemize}
    \item A link to a downloadable version of the dataset or simulation environment:

>>>>

This paper gathers all the supplementary material and goes as follows: \Cref{sec:proofs} details all the proofs of the main results. \Cref{sec:bftq-full} recalls the full scalable \BFTQ algorithm. \Cref{sec:lagragian} describes a naive alternative to \BFTQ based on Lagrangian Relaxation. The \Cref{sec:exp-supp} assembles all the assets for visualising and reproducing the experiments, including visualisations of policy executions, algorithms and environment parameters, and instructions for executing the attached source code. Finally we fill the Machine Learning Reproducibility Checklist and we justify each statement in \Cref{sec:ml-checklist}. Please note that all the materials referenced through web links are also available in the zip file of supplementary materials.

@@

\begin{proof}
Thanks to the introduction of the augmented spaces $\ocS, \ocA$ and dynamics $\ov{P}$, this proof is the same as that in classical multi-objective MDPs.

@@

\begin{align}
    Q_r^*(\os, \oa) &=  \max_{\rho, \pi'} Q_r^{\rho, \pi'}(\os', \oa') \label{eq:pthm_def}\\
    &= \max_{\rho, \pi'} R_r(s, a) + \gamma \sum_{\os'\in\ocS} P(\os' | \os, \oa) V_r^{\rho, \pi'}(\os') \label{eq:pthm_exp}\\
    &= R_r(s, a) + \gamma \sum_{\os'\in\ocS}  P(\os' | \os, \oa) \max_{\rho, \pi'} \sum_{\oa'\in\ocA} \rho(\oa' | \os')Q_r^{\pi'}(\os', \oa') \label{eq:pthm_marg}\\
    &= R_r(s, a) + \gamma \sum_{\os'\in\ocS}  P(\os' | \os, \oa) \max_\rho\sum_{\oa'\in\ocA}\rho(\oa' | \os')\max_{\pi'\in\Pi_a(\os')}Q_r^{\pi'}(\os', \oa') \label{eq:pthm_max}\\
    &= R_r(s, a) + \gamma \sum_{\os'\in\ocS}  P(\os' | \os, \oa) \max_\rho\expectedvalueover{\oa'\sim\rho}Q_r^*(\os', \oa') \label{eq:pthm_marg_def2}
\end{align}
where $\pi = (\rho, \pi')\in\Pi_a(\os)$ and $\pi'\in\Pi_a(\os')$.

This follows from:
\begin{enumerate}
\item[\eqref{eq:pthm_def}.] Definition of $Q^*$. 
\item[\eqref{eq:pthm_exp}.] Bellman Expectation expansion from \Cref{prop:bellman-expectation}.
\item[\eqref{eq:pthm_marg}.] Marginalisation on $\oa'$.
\item[\eqref{eq:pthm_max}.] \begin{itemize}
    \item Trivially $\max_{\pi'\in\Pi_a(\os')} \sum_{\oa'\in\ocA} \cdot \leq \sum_{\oa'\in\ocA} \max_{ \pi'\in\Pi_a(\os)} \cdot$
    \item Let $\ov{\pi}\in\argmax_{\pi'\in\Pi_a(\os')} Q_r^{\pi'}(\os', \oa')$, then:
    \begin{align*}
        \sum_{\oa'\in\ov{A}}\rho(\oa'|\os')\max_{\pi'\in\Pi_a(\os')}Q_r^{\pi'}(\os', \oa') &= \sum_{\oa'\in\ov{A}}\rho(\oa'|\os')Q_r^{\ov{\pi}}(\os', \oa') \\
        &\leq  \max_{\pi'\in\Pi_a(\os')} \sum_{\oa'\in\ov{A}}\rho(\oa'|\os')Q_r^{\pi'}(\os', \oa')
    \end{align*}
\end{itemize}
\item[\eqref{eq:pthm_marg_def2}.] Definition of $Q^*$.
\end{enumerate}

@@

\subsection{\Cref{rmk:contractivity-smooth}}
\label{proof:contraction-with-smooth}

\begin{remark*}
This proof makes use of insights detailed in the proof of \Cref{prop:bftq_pi_hull} (\Cref{sec:proof_pi_hull}), which we recommend the reader to consult first.
\end{remark*}

\begin{proof}
We now study the contractivity of $\cT$ when restricted to the functions of $\cL_\gamma$ defined as follows:
\begin{equation}
    \cL_\gamma = \left\{\begin{array}{cc}
         Q\in(\Real^2)^{\ocS\ocA}\text{ s.t. }\exists L<\frac{1}{\gamma}-1: \forall \os\in\ocS,\oa_1,\oa_2\in\ocA,   \\
         |Q_r(\os,\oa_1) - Q_r(\os,\oa_2)| \leq L|Q_c(\os,\oa_1) - Q_c(\os,\oa_2)|
    \end{array}\right\}
\end{equation}
That is, for all state $\os$, the set $Q(\os, \ocA)$ plot in the $(Q_c,Q_r)$ plane must be the \emph{graph} of a $L$-Lipschitz function, with $L<1/\gamma-1$.

@@

This proof makes use of insights detailed in the proof of \Cref{prop:bftq_pi_hull}, which we recommend the reader to consult first.

@@

%\section{Risk-Sensitive Exploration}
%\label{sec:risk-sensitive-supp}
%We recall the Risk-Sensitive Exploration in %\Cref{algo:risk-sensitive-exploration}:
%\input{source/risk-sensitive-explo-pseudo-code.tex}

@@

\subsection{Examples of different exploration strategies}
\label{subsec:exploration-examples}
We compare two approaches for constructing a batch of samples. The animations  \href{https://github.com/budgeted-rl/budgeted-rl.github.io\#risk-sensitive-exploration}{from this html page} display the trajectories collected in each intermediate sub-batch. The first row corresponds to a classical  risk neutral epsilon-greedy exploration policy while the second row showcases a risk-sensitive exploration strategy introduced in the paper. Each animation corresponds to a different seed.

@@

On the \texttt{highway-env} , the budgeted agents display a wide variety of behaviours. Animations are displayed \href{https://github.com/budgeted-rl/budgeted-rl.github.io#driving-styles}{on this html page}. When $\beta = 1$, the ego-vehicle drives in a very aggressive style: it immediately switches to the opposite lane and drives as fast as possible to pass slower vehicles, swiftly changing lanes to avoid incoming traffic. On the contrary when $\beta = 0$, the ego-vehicle is conservative: it stays on its lane and drives at a low velocity. With intermediate budgets such as $\beta = 0.2$, the agent sometimes decides to overtake its front vehicle but promptly steers back to its original lane afterwards.

@@

Animations are displayed  \href{https://github.com/budgeted-rl/budgeted-rl.github.io\#optimal-budgeted-policies-learnt-with-a-risk-sensitive-exploration}{on this html page} for the \texttt{corridors} environment. When the budget is low, the agent takes the safest path on the left. When the budget increases, it gradually switches to the other lane, earning higher rewards but also costs. This gradual process could not be achieved with a deterministic policy as it would chose either one path or the other. Each animation corresponds to a different seed.

@@

The following section displays environments and algorithms parameters and instructions to reproduce the exact same results displayed in \Cref{sec:experiments}.

@@

\subsubsection{Instructions for reproducibility}
\label{subsubsec:instruction-reproducibility}
To reproduce the result displayed in \Cref{sec:experiments}, first install the following conventional libraries for python3: pycairo, numpy, scipy and pytorch. Then, execute the commands in \Cref{fig:instructions} on a Linux Operating System. The Graphic Processing Unit used for experiments is an NVIDIA \texttt{GeForce GTX 1080 Ti} and the Computational Processing Unit is an Intel \texttt{Xeon E7}. Then, follow the instructions in \Cref{fig:instructions}. For more details, check this \href{https://github.com/budgeted-rl/budgeted-rl.github.io\#how-to-reproduce}{web address}.

@@

%\begin{minted}{sh}
\begin{lstlisting}[language=bash,caption={bash version}]
# Install highway-env
pip3 install --user git+https://github.com/eleurent/rl-agents
# clone the phd_code repository
git clone https://github.com/ncarrara/phd_code
# Change python path to the path of this repository
export PYTHONPATH="phd_code/ncarrara"
# Navigate to budgeted-rl folder
cd phd_code/ncarrara/budgeted-rl/
# Run main script using any config file
# Choose the range of seeds you want to test on
python3 main/egreedy/main-egreedy.py config/slot-filling.json 0 6
python3 main/egreedy/main-egreedy.py config/corridors.json 0 4
python3 main/egreedy/main-egreedy.py config/highway-easy.json 0 10
\end{lstlisting}
%\end{minted}

@@

    \begin{itemize}\item \textbf{yes}, see \Cref{sec:intro}, \Cref{sec:bdp}, \Cref{sec:brl},
    \Cref{sec:exploration} and \Cref{sec:bftq-full}.
    \end{itemize}
    \item An analysis of the complexity (time, space, sample size) of any algorithm:
    \begin{itemize}\item \textbf{yes}, see \Cref{subsec:parallel-computing}.
    \end{itemize}
    \item A link to a downloadable source code, with specification of all dependencies, including external libraries:
    \begin{itemize}\item \textbf{yes}, see \Cref{subsubsec:instruction-reproducibility} and the folder \texttt{code} in the supplementary material zip file.
    \end{itemize}
\end{itemize}

@@

\begin{itemize}
    \item  A statement of the result:
     \begin{itemize}
        \item  \textbf{yes}, see \Cref{sec:bdp} and \Cref{sec:brl}.\end{itemize}
    \item A clear explanation of any assumptions:
    \begin{itemize}
        \item  we make one assumption in \Cref{sec:bdp}. We assume the program is feasible for any state. If not, no algorithm would be able to solve it anyway.\end{itemize}
    \item A complete proof of the claim:
    \begin{itemize}
        \item  \textbf{yes}, see \Cref{sec:proofs}. 
    \end{itemize}
\end{itemize}

@@

        \begin{itemize}
            \item\textbf{yes}, see \Cref{sec:experiments} and \Cref{sec:algorithms-parameters}.
        \end{itemize}
    \item A link to a downloadable version of the dataset or simulation environment:
